---
title: "Human Activity Recognition - Predicting how well weight lifting excercises were performed using sensory data"
subtitle: "Practical Machine Learning Final Assignment"
author: "atomasovszky"
date: "4/27/2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = FALSE, 
    messages = FALSE,
    dpi = 300,
    out.width = "350pt",
    fig.align = "center"
)

source("libraries_final_assignment.R")

train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train <- fread(train_url)
test  <- fread(test_url)
```


## Synopsis

The goal of this work is to use on-body sensor data to predict whether a particular excercise is performed correctly.
The 6 participants of the study our data is coming from were equippped with movement sensors on their arm, forearm, a belt sensor on their waist and one on their dumbell, and they've been asked to perform weight lifting in a proper way (classe = A), and with some common mistakes (classe = {B, C, D, E}).


&nbsp;
&nbsp;


## Feature selection

Besides the obvious ones (ids, timestamps, etc.) we also left summary variables out as they have missing values for most of the records.

```{r feature_selection}
summary_vars <- grep(
    pattern = "^kurtosis|^skewness|^max|^min|^amplitude|^var|^avg|^stddev", 
    x       = names(train),
    value   = TRUE
)

variables_to_omit <- c(
    summary_vars,
    grep("window|V1|user|timestamp", names(train), value = TRUE)
)

train_clean <- train[, .SD, .SDcols = setdiff(names(train), variables_to_omit)]
```


&nbsp;
&nbsp;


## Checking the distribution of features

The three boxplots included here shows that there are a few outliers in the dataset.
These come from two cases, id 5373, and 9274. 
We omitted these during the training.

```{r feature_distribution_check}
walk(setdiff(names(train_clean), "classe"), ~{
    filename <- glue("figures/{.x}.png")
    if (!file.exists(filename)) {
        ggplot(train_clean, aes_string("classe", .x, fill = "classe")) +
            geom_boxplot(alpha = .75)
        ggsave(filename)
    }
})

ggplot(train_clean, aes(classe, magnet_dumbbell_y, fill = classe)) +
    geom_boxplot(alpha = .75)

ggplot(train_clean, aes(classe, gyros_dumbbell_x, fill = classe)) +
    geom_boxplot(alpha = .75)

ggplot(train_clean, aes(classe, gyros_forearm_x, fill = classe)) +
    geom_boxplot(alpha = .75)

outliers <- train[magnet_dumbbell_y < -2000 | gyros_forearm_x < -20][, V1]
outliers
train_clean <- train_clean[magnet_dumbbell_y > -2000 | gyros_forearm_x > -20]
```
